### 2018-10-22 src/4_OnlineLearning.tex
correccion despues del examen

\section{Introduction}
Machine Learning (ML) studies computer algorithms for learning something such as
to complete a task, to make accurate predictions or to behave intelligently. The
learning is always based on samples and the objective is about to do better in
the future based on the past experiences in an automatic way. ML is a sub-area
of artificial intelligence and broadly intersects with other fields such as
statistics, mathematics, physics and computer science.  There are many examples
of machine learning problems: time series forecasting, image processing, face
detection, spam filtering, weather prediction, search engines, among many
others.

@@

\begin{description}
\item[Supervised learning]  the input data is a tuple which contains the example
input and its desired outputs (also called labels). The list of tuples is called
the training set. The concept of supervised learning comes from the supervisor,
acting as a teacher in the learning process. The goal is to learn a general rule
that maps inputs to outputs optimising a target function. There are two related
problem types in supervised learning: classification and regression problems
\cite{bishop2006}. Its two mainstream approaches are: support vector machines
(SVMs) \cite{vapnik1998} and ensemble learning \cite{breiman1998}. Furthermore,
supervised learning can be categorised into offline or batch learning and online
learning (see section \ref{sec:onoffline}). 
\item[Unsupervised learning] also known as clustering \cite{ben2005}. In this
type of learning no labels are given and the system has to find a structure on
its own, discovering hidden patterns in data.
\item[Reinforcement learning] is the problem faced by an agent that must learn
through trial-and-error interactions with a dynamic environment. It is based on
programming agents by reward and punishment without the need to specify how the
task is to be achieved \cite{sutton1998}.
\end{description}

@@

%Supervised learning consists in finding a learning function $f: X \rightarrow Y$
%which models a set of examples, called training set $S$, which have been drawn
%randomly and independently according to a unknown joint distribution function
%$p(x,y)$ on $X \times Y$:
%
%
%$$
%S = \{ (x_k,y_k)\in X\times Y : k=1,...,n \}
%$$
%
%\noindent where $X \subseteq \mathbb{R}^m$ and $Y = \mathbb{R}$ in case of
%regression problems and $Y = \{1,-1\}$  if it is a classification problem. 
%
%
%A learning algorithm $\mathcal{A}$ takes as input a data set $S \in
%\mathscr{T}=\mathscr{P}(X\times Y)$ and selects from $\mathscr{H}$ a function
%$f$ such that $f(\mathbf{x})\approx y $ in a predictive way:
%
%\begin{eqnarray*}
%\mathcal{A}: \mathscr{T} &\rightarrow & \mathscr{H} \\
%S &\rightarrow &\mathcal{A}(S) = f
%\end{eqnarray*}
%
%\noindent where $\mathscr{H}$ is called the {\em hypothesis space}, i.e. the space
%of functions that the algorithm is allowed to search.

%------------------------DRAFT PROFE -------------------------------
The basic setting for supervised learning is a {\em data set\/} $X$, 
and an  {\em outcome set\/} $Y$.
For our purposes $X$ will be $\R^m$, $m\in\N$, and $Y$ either $\R$
in the case of regression problems, or $\{-1,1\}$ in the case of
classification problems.
In addition we have a plurality of {\em training sets\/}
$$
S=\bigg\{ (\mathbf{x}_k,y_k)\in X\times Y \mid k=1,\dots,n \bigg\}
\subseteq X\times Y\,,
$$
where the members of $S$ have been drawn randomly and independently
from $X\times Y$ according to an {\em unknown\/} joint distribution
function $p(\mathbf{x},y)$ on $X\times Y$.
We gather all these training sets in a subset $\mathscr{T}$ of
$\mathscr{P}(X\times Y)$.

@@

%------------------------FIN DRAFT PROFE -------------------------------
{\em Supervised learning\/} consists then in finding a {\em learning function\/}
$f:X\to Y$ that minimise the expected error of the loss function $V(f(\mathbf{x}),y): Y
\times Y$ defined as:

%\begin{eqnarray*}
%\label{eq:lossfunction}
% V & \rightarrow & \mathbb{R}^+ \\
%  (f(\mathbf{x}), y) &\rightarrow & V(f(\mathbf{x}),y)
%\end{eqnarray*}

@@

\begin{equation}
\label{eq:expetedrisk}
%E[V(f(\mathbf{x}),y)] = \int_{X\times Y} V(f(\mathbf{x}),y)\;dp(x,y)\,,
E\big[ V( f(\cdot),\cdot) \big]
= \int_{X\times Y} V\big( f(\mathbf{x},y \big)\,dp(\mathbf{x},y)\,.
\end{equation}

For regression, the most common loss function is square loss or L2 loss function:
\begin{equation}
\label{eq:l2loss}
V(f(\mathbf{x}),y) = (f(\mathbf{x})-y)^2
\end{equation}
\noindent another option is the L1 loss
\begin{equation}
\label{eq:l1loss}
V(f(\mathbf{x}),y) = |f(\mathbf{x})-y| \, .
\end{equation}
\noindent the choice of loss function here gives rise to several well-known
learning algorithms such as regularised least squares and support vector
machines. %verificar esto

@@

The ERM is a Riemann sum approximation of the expected error of $f$ shown in
equation \ref{eq:expetedrisk}. By the law of large numbers it is known that:

\begin{equation}
%\frac{1}{n} \sum_{i=1}^n V(f(x_i),y_i) \xrightarrow{n \rightarrow \infty}  E[V(f(\mathbf{x}),y)]
\frac1n\,\sum_{i=1}^n V\big( f(\bfx_i,y_i \big)
\overset{n\to\infty}{\longrightarrow} E\big[ V(f(\cdot),\cdot)\big]\,.
\end{equation}

@@

\item[Generalisation accuracy on new examples] is the capability of the
algorithm to generate the right output for an input instance outside the
training set. It is measured by the expected risk (equation \ref{eq:expetedrisk}).
\item[The amount of training data] in most cases, generalisation accuracy
increases as the amount of training data increases.
\end{description}

@@

Models learning error can be split into two main components: error due to bias
and error due to variance. Bias measures the prediction error of the model
$\hat{f}$ from the real value, see equation \ref{eq:biastrade}. Variance is
taken as the variability of a model prediction for a given data point or its
sensitivity to small fluctuations in the input data, see equation
\ref{eq:variancetrade}.

If we have a learning model $\hat{f}$, its expected generalisation error on a
testing sample $x$ can be decomposed as shown in equation \ref{eq:bvtrade} (proof
in \cite{geman1992neural}):

@@

Figure \ref{fig:bvtradeoff} shows how
testing error can be decomposed into bias and variance components. 
The bias shown in equation \ref{eq:biastrade} can be obtain through resampling
of the input data $x$, therefore an average of the predicted values can be
calculated. If these prediction values are substantially different to the true
value, the bias will be high.

@@

\noindent where $X \subseteq \mathbb{R}^m$ and $Y = \mathbb{R}$ , $V$ is the
loss function defined in the equation \ref{eq:lossfunction} and $\mathcal{R}(f)$
is the regulariser, a penalisation on $f$. One example of regularisation in
linear models is Ridge Regression (RR), which is a regularised least squares
method.

@@

To solve equation \ref{eq:lsm} is equivalent to find a solution of:
\begin{equation}
\label{eq:regproblem}
\underset{m \times n}{\mathbf{A}} \enskip \underset{n \times
l}{\mathbf{\mathbf{X}}} = \underset{m \times l}{\mathbf{Y}}
\end{equation}

The optimal solution for $\mathbf{\mathbf{X}}$ is:
\begin{equation}
\label{eq:MP}
\mathbf{\mathbf{X}}= \mathbf{A}^+ \mathbf{Y}
\end{equation}

@@

%\textbf{Demo}\quad
%
%To solve equation~(\ref{eq:pseudoinverse}) is equivalent to solve the following
%optimisation problem:
%
%
%\begin{equation}
%\label{eq:RRproblem2}
%\underset{\mathbf{X}(\lambda)}{\text{min}} \quad \|
%\mathbf{C}\mathbf{\mathbf{X}} - \mathbf{F} \|_2^2
%\end{equation}
%
%\noindent where
%
%\begin{equation*}
%	\mathbf{C}=
%	\begin{bmatrix} \quad &\mathbf{A}& \quad \\ \hdashline \quad
%& \sqrt\lambda\mathbb{I}&\quad  
%\end{bmatrix} 
%\quad \text{and} \quad
%	\mathbf{F}=\begin{bmatrix} \quad &\mathbf{Y}& \quad \\ \hdashline \quad
%& \mathbf{0}&\quad  \end{bmatrix} \, .
%\end{equation*}
%
%Applying equation~(\ref{eq:MP}) and considering that $\mathbf{C}^\top
%\mathbf{C} = \mathbf{A}^\top \mathbf{A} + \lambda \mathbb{I}$ and 
%$\mathbf{C}^\top \mathbf{F}=\mathbf{A}^\top \mathbf{Y} $ we have:
%
%\begin{eqnarray*}
%\mathbf{X}(\lambda)&=&(\mathbf{C}^\top
%\mathbf{C})^{-1}\mathbf{C}^\top \mathbf{F} \\
%&=& (\mathbf{A}^\top \mathbf{A} + \lambda \mathbb{I})^{-1} \mathbf{A}^\top \mathbf{Y}
%\end{eqnarray*}
%
%$\blacksquare$

@@

\begin{equation}
    \label{eq:compactsvd}
    \underset{m \times n}{\mathbf{A}}=
    \underset{m \times k}{\mathbf{U_1}} \enskip
    \underset{k \times k}{\mathbf{\Sigma_1}} \enskip
    \underset{k \times n}{\mathbf{V_1}^\top}
\end{equation}

@@

\begin{equation}
\label{eq:pseudoinversesvd}
\mathbf{A}^+ = \mathbf{V_1\Sigma_1^{-1}U_1^\top}
\end{equation}

@@

\begin{equation}
\label{eq:RRproblem} 
J(\mathbf{\mathbf{X}}) =  \| \mathbf{A}\mathbf{\mathbf{X}} - \mathbf{Y} \|_2^2  + \lambda
 \| \mathbf{\mathbf{X}}\| ^2
\end{equation}

@@

\begin{equation}
\label{eq:optsolRR}
\mathbf{\mathbf{X}}_*=(\mathbf{A}^\top \mathbf{A}+\lambda \mathbb{I})^{-1}\mathbf{A}^\top y \, ,
\end{equation}

@@

where $\sigma_1 \geq \sigma_2 \geq \dots \geq \sigma_n$.

@@

If matrix $\mathbf{A}$ is non-singular, its condition number can be
expressed in terms of its singular values. The effect of adding the
regularization term affects the condition number as follows:

\begin{eqnarray*}
\kappa_{ols} &=& \|\mathbf{A}\| \|\mathbf{A}^{-1}\|=\frac{\sigma_1}{\sigma_n} \\
\kappa_{ridge} &=& \|\mathbf{A}^\intercal \mathbf{A} + \lambda \mathbb{I}\| 
\|(\mathbf{A}^\intercal \mathbf{A} + \lambda \mathbb{I})^{-1}\|=\frac{\sigma_1+\lambda}{\sigma_n + \lambda} \,
\end{eqnarray*}

@@

One of the ways to determine parameter $\lambda$ is using the bias-variance tradeoff (see section \ref{sec:biasvar}). This parameter is crucial for ridge regression since it could
reduce the expected prediction error by reducing variance, considering a biased
estimator. 
It is known that the prediction error can be express as a decomposition between bias and variance.

@@

Incremental learning is very useful when there is no need to record fundamental
data and only a summary needs to be retained. Due to this, incremental
algorithms are often characterised as memoryless, because no memory of past data
is required.  The algorithm is online but not incremental if it doesn't produce
the same result for all observations that the corresponding batch algorithm
would for these same observations.

Algorithm~\ref{alg:onlinealg} shows the online learning algorithm structure:

\begin{algorithm}[ht]
\begin{algorithmic}[1]
    \STATE Receives input $\mathbf{x}_t$
    \STATE Makes prediction $\mathbf{\hat{y}}_t$
    \STATE Receives response $\mathbf{y}_t$
    \STATE Incurs loss $l_t(\mathbf{y}_t,\mathbf{\hat{y}}_t)$
\end{algorithmic}
\caption{Structure of a Learning System}
\label{alg:onlinealg}
\end{algorithm}

\noindent where $l$ is some loss function. Performance is later measured after
$T$ trials as:

\begin{equation*}
L_T = \sum_{t=1}^T l_t(\mathbf{y}_t,\mathbf{\hat{y}}_t)
\end{equation*}

The objective is to minimise this loss function for all instances.
The quality of online learning algorithms is measured by a quantity known as
regret which is the difference between the performance of the online algorithm
and its optimal predictor $E^* \in \Theta$ given by:

@@

The AAR, proposed by~\cite{vovk2001} (also known as Vovk-Azoury-Warmuth predictor ~\cite{azoury2001}), is an application of the aggregating
algorithm to the problem of regression. The idea is introduce the new input
vector $\mathbf{x}_{m+1}$ to solve the model parameters:

\begin{equation}
\label{eq:AARexpand}
\mathbf{X}_{aar} = \displaystyle \big (\sum_{t=1}^{m+1}
\mathbf{a}_t \mathbf{a}_t  ^\intercal + \gamma \mathbb{I}\big )^{-1}
\sum_{t=1}^m \mathbf{a}_t \mathbf{y}_t \, .
\end{equation}

>>>>

\section{Introduction}
Machine Learning (ML) studies computer algorithms for learning patterns to complete a task, to make accurate predictions or to behave intelligently. The
learning is always based on samples and the objective is about to do better in
the future based on the past experiences in an automatic way. ML is a sub-area
of artificial intelligence and broadly intersects with other fields such as
statistics, mathematics, physics and computer science.  There are many examples
of machine learning problems: time series forecasting, image processing, face
detection, spam filtering, weather prediction, search engines, among many
others.

@@

\begin{description}
\item[Supervised learning]  the input data is a tuple which contains the example
input and its desired outputs (also called labels). The list of tuples is called
the training set. The concept of supervised learning comes from the supervisor,
acting as a teacher in the learning process. The goal is to learn a general rule
that maps inputs to outputs optimising a target function. There are two related
problem types in supervised learning: classification and regression problems
\cite{bishop2006}. Its two mainstream approaches are: support vector machines
(SVMs) \cite{vapnik1998} and ensemble learning \cite{breiman1998}. Furthermore,
supervised learning can be categorised into offline or batch learning and online
learning (see section~\ref{sec:onoffline}). 
\item[Unsupervised learning] also known as clustering \cite{ben2005}. In this
type of learning no labels are given and the system has to find a structure on
its own, discovering hidden patterns in data.
\item[Reinforcement learning] is the problem faced by an agent that must learn
through trial-and-error interactions with a dynamic environment. It is based on
programming agents by reward and punishment without the need to specify how the
task is to be achieved \cite{sutton1998}.
\end{description}

@@

The basic setting for supervised learning is a {\em data set\/} $X$, 
and an  {\em outcome set\/} $Y$.
For our purposes $X$ will be $\R^m$, $m\in\N$, and $Y$ either $\R$
in the case of regression problems, or $\{-1,1\}$ in the case of
classification problems.
In addition we have a plurality of {\em training sets\/}
$$
S=\bigg\{ (\mathbf{x}_k,y_k)\in X\times Y \mid k=1,\dots,n \bigg\}
\subseteq X\times Y\,,
$$
where the members of $S$ have been drawn randomly and independently
from $X\times Y$ according to an {\em unknown\/} joint distribution
function $p(\mathbf{x},y)$ on $X\times Y$.
We gather all these training sets in a subset $\mathscr{T}$ of
$\mathscr{P}(X\times Y)$.

@@

{\em Supervised learning\/} consists then in finding a {\em learning function\/}
$f:X\to Y$ that minimise the expected error of the loss function $V(f(\mathbf{x}),y): Y
\times Y$ defined as:

@@

\begin{equation}
\label{eq:expetedrisk}
E\big[ V( f(\cdot),\cdot) \big]
= \int_{X\times Y} V\big( f(\mathbf{x},y) \big)\,dp(\mathbf{x},y)\,.
\end{equation}

For regression, the most common loss function is square loss or $L_{2}$ loss function:
\begin{equation}
\label{eq:l2loss}
V(f(\mathbf{x}),y) = (f(\mathbf{x})-y)^2
\end{equation}
\noindent another option is the L1 loss
\begin{equation}
\label{eq:l1loss}
V(f(\mathbf{x}),y) = |f(\mathbf{x})-y| \, .
\end{equation}
\noindent the choice of loss function here gives rise to several well-known
learning algorithms such as regularised least squares and support vector
machines. %verificar esto

@@

The ERM is a Riemann sum approximation of the expected error of $f$ shown in
equation~\eqref{eq:expetedrisk}. By the law of large numbers it is known that:

\begin{equation}
\frac1n\,\sum_{i=1}^n V\big( f(\bfx_i,y_i) \big)
\overset{n\to\infty}{\longrightarrow} E\big[ V(f(\cdot),\cdot)\big]\,.
\end{equation}

@@

\item[Generalisation accuracy on new examples] is the capability of the
algorithm to generate the right output for an input instance outside the
training set. It is measured by the expected risk (equation~\eqref{eq:expetedrisk}).
\item[The amount of training data] in most cases, generalisation accuracy
increases as the amount of training data increases.
\end{description}

@@

Models learning error can be split into two main components: error due to bias
and error due to variance. Bias measures the prediction error of the model
$\hat{f}$ from the real value, see equation~\eqref{eq:biastrade}. Variance is
taken as the variability of a model prediction for a given data point or its
sensitivity to small fluctuations in the input data, see equation
\ref{eq:variancetrade}.

If we have a learning model $\hat{f}$, its expected generalisation error on a
testing sample $x$ can be decomposed as shown in equation~\eqref{eq:bvtrade} (proof
in \cite{geman1992neural}):

@@

Figure \ref{fig:bvtradeoff} shows how
testing error can be decomposed into bias and variance components. 
The bias shown in equation~\eqref{eq:biastrade} can be obtain through resampling
of the input data $x$, therefore an average of the predicted values can be
calculated. If these prediction values are substantially different to the true
value, the bias will be high.

@@

\noindent where $X \subseteq \mathbb{R}^m$ and $Y = \mathbb{R}$ , $V$ is the
loss function defined in the equation~\eqref{eq:lossfunction} and $\mathcal{R}(f)$
is the regulariser, a penalisation on $f$. One example of regularisation in
linear models is Ridge Regression (RR), which is a regularised least squares
method.

@@

To solve equation~\eqref{eq:lsm} is equivalent to find a solution of:
\begin{equation}
\label{eq:regproblem}
\underset{m \times n}{\mathbf{A}} \enskip \underset{n \times
l}{\mathbf{\mathbf{X}}} = \underset{m \times l}{\mathbf{Y}} \;.
\end{equation}

The optimal solution for $\mathbf{\mathbf{X}}$ is:
\begin{equation}
\label{eq:MP}
\mathbf{\mathbf{X}}= \mathbf{A}^+ \mathbf{Y} \;,
\end{equation}

@@

@@

\begin{equation}
    \label{eq:compactsvd}
    \underset{m \times n}{\mathbf{A}}=
    \underset{m \times k}{\mathbf{U_1}} \enskip
    \underset{k \times k}{\mathbf{\Sigma_1}} \enskip
    \underset{k \times n}{\mathbf{V_1}^\top} \;,
\end{equation}

@@

\begin{equation}
\label{eq:pseudoinversesvd}
\mathbf{A}^+ = \mathbf{V_1\Sigma_1^{-1}U_1^\top} \;.
\end{equation}

@@

\begin{equation}
\label{eq:RRproblem} 
J(\mathbf{\mathbf{X}}) =  \| \mathbf{A}\mathbf{\mathbf{X}} - \mathbf{Y} \|_2^2  + \lambda
 \| \mathbf{\mathbf{X}}\| ^2 \;,
\end{equation}

@@

\begin{equation}
\label{eq:optsolRR}
\mathbf{\mathbf{X}}_*=(\mathbf{A}^\top \mathbf{A}+\lambda \mathbb{I})^{-1}\mathbf{A}^\top y \;.
\end{equation}

@@

\noindent where $\sigma_1 \geq \sigma_2 \geq \dots \geq \sigma_n$.

@@

If matrix $\mathbf{A}$ is non-singular, its condition number can be
expressed in terms of its singular values. The effect of adding the
regularization term affects the condition number as follows:
\begin{eqnarray*}
\kappa_{ols} &=& \|\mathbf{A}\| \|\mathbf{A}^{-1}\|=\frac{\sigma_1}{\sigma_n} \\
\kappa_{ridge} &=& \|\mathbf{A}^\intercal \mathbf{A} + \lambda \mathbb{I}\| 
\|(\mathbf{A}^\intercal \mathbf{A} + \lambda \mathbb{I})^{-1}\|=\frac{\sigma_1+\lambda}{\sigma_n + \lambda} \;.
\end{eqnarray*}

@@

One of the ways to determine parameter $\lambda$ is using the bias-variance tradeoff (see section~\ref{sec:biasvar}). This parameter is crucial for ridge regression since it could
reduce the expected prediction error by reducing variance, considering a biased
estimator. 
It is known that the prediction error can be express as a decomposition between bias and variance.

@@

Incremental learning is very useful when there is no need to record fundamental
data and only a summary needs to be retained. Due to this, incremental
algorithms are often characterised as memoryless, because no memory of past data
is required.  The algorithm is online but not incremental if it doesn't produce
the same result for all observations that the corresponding batch algorithm
would for these same observations.
\newpage
Algorithm~\ref{alg:onlinealg} shows the online learning algorithm structure:
\begin{algorithm}[ht]
\begin{algorithmic}[1]
    \STATE Receives input $\mathbf{x}_t$
    \STATE Makes prediction $\mathbf{\hat{y}}_t$
    \STATE Receives response $\mathbf{y}_t$
    \STATE Incurs loss $l_t(\mathbf{y}_t,\mathbf{\hat{y}}_t)$
\end{algorithmic}
\caption{Structure of a Learning System}
\label{alg:onlinealg}
\end{algorithm}
\noindent where $l$ is some loss function. Performance is later measured after
$T$ trials as:
\begin{equation*}
L_T = \sum_{t=1}^T l_t(\mathbf{y}_t,\mathbf{\hat{y}}_t)
\end{equation*}
The objective is to minimise this loss function for all instances.
The quality of online learning algorithms is measured by a quantity known as
regret which is the difference between the performance of the online algorithm
and its optimal predictor $E^* \in \Theta$ given by:

@@

The AAR, proposed by Vovk~\cite{vovk2001} (also known as Vovk-Azoury-Warmuth predictor ~\cite{azoury2001}), is an application of the aggregating
algorithm to the problem of regression. The idea is introduce the new input
vector $\mathbf{x}_{m+1}$ to solve the model parameters:

\begin{equation}
\label{eq:AARexpand}
\mathbf{X}_{aar} = \displaystyle \left(\sum_{t=1}^{m+1}
\mathbf{a}_t \mathbf{a}_t  ^\intercal + \gamma \mathbb{I} \right) ^{-1}
\sum_{t=1}^m \mathbf{a}_t \mathbf{y}_t \, .
\end{equation}

### 2018-10-22 src/5_1_Proposal.tex
correccion despues del examen

The AVECM algorithm is presented in section \ref{sec:51methodology}.
In section \ref{sec:51results} we describe the tests carried on to assess the
accuracy and the execution time of AVECM.  This section also includes a
description of the test data.  Section \ref{sec:51conclusions} contains the
conclusions and a discussion of future research.

@@

Figure \ref{fig:hists} shows the distribution of the number of cointegration
vectors given by the Johansen method for different values of $L=[2,6,10,14]$
hours and $p=1$ . This procedure was carried out by a sliding window of
historical data moving 1000 times. We observed that the distribution of
cointegration vectors changed with different values of $L$. When
$L=2$ hours, there was no cointegration in more than 60\% of the iterations.
Cointegration increased when $L=6$ hours and was maximum when $L=10$ hours where
one cointegration vectors was found for all 1000 iterations. The ocurrence of
cointegration started to decrease at $L=14$ hours.

From section \ref{sec:vecm} we know that $r=0$ means no cointegration and
$r=l$ (we are using four rates, so $l=4$) reveals that no process is I(1) but
stationary.  The interesting cases of cointegration are those where $r$ lies
strictly between $0$ and $4$, i.e. $0<r<4$.

@@

The goal of our next experiment was to find a relationship between the ratio $PC$
and the performance measure MSE (see equation \ref{eq:MSE}). $L$ was defined
between $[2,14]$ hours, that corresponded to $[720, 5040]$ data points, and $p$
took values between $[1,5]$.

@@

Execution time depends directly on $L$, $p$ and $nparams$ used, since they
determine the size of matrix $\mathbf{A}$ (see equation \ref{eq:vecA}) and
therefore affect the OLS function execution time.  Therefore, if we try more
combinations of $L$ and $p$ (increasing $nparams$) the serial algorithm will
take longer. For this financial time series we are interested in execution times
below 10 seconds  (the time series frequency).

@@

The superior figure in \ref{fig:extimes} shows that best performance accuracy
measurements are achieved in times near or below 10 seconds in the parallel
version. Contrarily, serial times are higher, above 10 seconds in most cases.
Figure~\ref{fig:extimes} also shows the speed-up in computing time for AVECM
which is near 9X if we use more than 25 parameters ($nparams$).

>>>>

The AVECM algorithm is presented in section~\ref{sec:51methodology}.
In section~\ref{sec:51results} we describe the tests carried on to assess the
accuracy and the execution time of AVECM.  This section also includes a
description of the test data.  Section~\ref{sec:51conclusions} contains the
conclusions and a discussion of future research.

@@

Figure~\ref{fig:hists} shows the distribution of the number of cointegration
vectors given by the Johansen method for different values of $L=[2,6,10,14]$
hours and $p=1$ . This procedure was carried out by a sliding window of
historical data moving 1000 times. We observed that the distribution of
cointegration vectors changed with different values of $L$. When
$L=2$ hours, there was no cointegration in more than 60\% of the iterations.
Cointegration increased when $L=6$ hours and was maximum when $L=10$ hours where
one cointegration vectors was found for all 1000 iterations. The ocurrence of
cointegration started to decrease at $L=14$ hours.

From section~\ref{sec:vecm} we know that $r=0$ means no cointegration and
$r=l$ (we are using four rates, so $l=4$) reveals that no process is I(1) but
stationary.  The interesting cases of cointegration are those where $r$ lies
strictly between $0$ and $4$, i.e. $0<r<4$.

@@

The goal of our next experiment was to find a relationship between the ratio $PC$
and the performance measure MSE (see equation~\ref{eq:MSE}). $L$ was defined
between $[2,14]$ hours, that corresponded to $[720, 5040]$ data points, and $p$
took values between $[1,5]$.

@@

Execution time depends directly on $L$, $p$ and $nparams$ used, since they
determine the size of matrix $\mathbf{A}$ (see equation~\ref{eq:vecA}) and
therefore affect the OLS function execution time.  Therefore, if we try more
combinations of $L$ and $p$ (increasing $nparams$) the serial algorithm will
take longer. For this financial time series we are interested in execution times
below 10 seconds  (the time series frequency).

@@

The superior figure in~\ref{fig:extimes} shows that best performance accuracy
measurements are achieved in times near or below 10 seconds in the parallel
version. Contrarily, serial times are higher, above 10 seconds in most cases.
Figure~\ref{fig:extimes} also shows the speed-up in computing time for AVECM
which is near 9X if we use more than 25 parameters ($nparams$).

### 2018-10-22 src/5_2_Proposal.tex
first draft ready to print
correccion despues del examen

\begin{itemize}
\item The function \texttt{getJohansen} returns cointegration vectors given by
the Johansen method considering the trace statistic test at 95\%
significance level.
\item The function \texttt{vecMatrix} returns the matrices~(\ref{eq:vecA})
and~(\ref{eq:vecY}) that allows VECM to be solved.
\item The function \texttt{vecMatrixOnline} returns the
matrices~(\ref{eq:vecA}) and~(\ref{eq:vecY}) aggregating new data and removing
the old one, avoiding calculation of the matrix $\mathbf{A}$ from scratch.
\item Out-of-sample outputs are saved in the variables 
$\mathbf{Y}_{\text{true}}$ and $\mathbf{Y}_{\text{pred}}$.
\item The model is solved using OLS.
\item In-sample outputs are saved in the variables $\Delta
\mathbf{y}_{\text{true}}$ and $\Delta \mathbf{y}_{\text{pred}}$.
\item The function \texttt{mape} gets the in-sample MAPE for the $l$ time
series.
\item Cointegration vectors are obtained at the beginning and when they are required to be updated. This updating is decided based on the in-sample MAPE of the last $n$ inputs. The average of all
MAPEs are stored in the variable $e$. If the average of MAPEs
($\texttt{mean}(e)$) is above a certain error given by the mean\_error threshold, cointegration vectors are updated.
\item If new cointegration vectors are required, the function
\texttt{vecMatrixUpdate} only updates the corresponding columns of matrix
$\mathbf{A}$ affected by those vectors (see equation~\ref{eq:vecA}).
\end{itemize}
\newpage

A pseudocode of the algorithm \ref{alg:52proposal} is detailed in the figure\ref{fig:pseudocodigo}.
\newline

@@

    \draw [arrow] (dec2) --  node[anchor=east] {no} (dec3);
    \draw [arrow] (dec3.west) -| node[anchor=east] {no} (pro5);
    \draw [arrow] (pro5) |- (pro1);
    \draw [arrow] (dec3) --  node[anchor=east] {si} (out1);
    %\draw [arrow] (out1) -- (stop);
    \end{tikzpicture}
}

>>>>

\begin{itemize}
\item The function \texttt{getJohansen} returns cointegration vectors given by
the Johansen method considering the trace statistic test at 95\%
significance level.
\item The function \texttt{vecMatrix} returns the matrices~\eqref{eq:vecA}
and~\eqref{eq:vecY} that allows VECM to be solved.
\item The function \texttt{vecMatrixOnline} returns the
matrices~\eqref{eq:vecA} and~\eqref{eq:vecY} aggregating new data and removing
the old one, avoiding calculation of the matrix $\mathbf{A}$ from scratch.
\item Out-of-sample outputs are saved in the variables 
$\mathbf{Y}_{\text{true}}$ and $\mathbf{Y}_{\text{pred}}$.
\item The model is solved using OLS.
\item In-sample outputs are saved in the variables $\Delta
\mathbf{y}_{\text{true}}$ and $\Delta \mathbf{y}_{\text{pred}}$.
\item The function \texttt{mape} gets the in-sample MAPE for the $l$ time
series.
\item Cointegration vectors are obtained at the beginning and when they are required to be updated. This updating is decided based on the in-sample MAPE of the last $n$ inputs. The average of all
MAPEs are stored in the variable $e$. If the average of MAPEs
($\texttt{mean}(e)$) is above a certain error given by the mean\_error threshold, cointegration vectors are updated.
\item If new cointegration vectors are required, the function
\texttt{vecMatrixUpdate} only updates the corresponding columns of matrix
$\mathbf{A}$ affected by those vectors (see equation~\eqref{eq:vecA}).
\end{itemize}

@@

    \draw [arrow] (dec2) --  node[anchor=east] {no} (dec3);
    \draw [arrow] (dec3.west) -| node[anchor=east] {no} (pro5);
    \draw [arrow] (pro5) |- (pro1);
    \draw [arrow] (dec3) --  node[anchor=east] {si} (out1);
    %\draw [arrow] (out1) -- (stop);
    \end{tikzpicture}
}
A pseudocode of the algorithm \ref{alg:52proposal} is detailed in the figure\ref{fig:pseudocodigo}.

### 2018-10-22 src/5_3_Proposal.tex
inicio revision capitulo 3x
correccion despues del examen

Since cointegration vectors represent a long-run relationship between the time
series, they vary little in time. This proposal also includes to obtain new
cointegration vectors only when this long-term relationship changes. This change
is detected by tracking the Mean Absolute Percentage Error (MAPE) defined in the
equation \ref{eq:MAPE}, of the last $n$ in-sample forecasts.

@@

\noindent where matrix $\mathbf{B},\mathbf{X},\mathbf{E}$ and $\mathbf{A}$ were
previously defined in equations \ref{eq:vecY} , \ref{eq:vecX} , \ref{eq:vecE} and
\ref{eq:vecA} respectively.

@@

The RR solution $\mathbf{\hat{X}}(t)$ of equation \ref{eq:vareqt1} using the
sliding window matrices defined above is:

@@

\begin{itemize}
\item The function \texttt{getJohansen} returns cointegration vectors given by the
Johansen method considering the trace statistic test at 95\% level of
significance. This procedure is called at the first step of the algorithm and
every time in-sample MAPE ($\mathbf{e}$) exceeds threshold error defined (mean\_error).
\item The function \texttt{vecMatrix} returns VECM
matrices $\mathbf{A}(t),\mathbf{B}(t)$ shown in equation~(\ref{eq:notation}). This
method is only required at the first step.
\item The function \texttt{vecMatrixOnline} returns new rows $\mathbf{a}_t^\top$ and
$\mathbf{b}_t^\top$ given new input data $\mathbf{y}_t$.
\item The Solver class (see algorithm \ref{alg:solver}) obtain VECM parameters  
using RR, AAR or OLS methods.
\item The function \texttt{vecMatrixUpdate} updates matrix $\mathbf{A}$ when new
cointegration vectors are required (matrix $\mathbf{B}$ is not affected). Model
parameters matrix $\mathbf{X}$ is also updated later.
\item The number of cointegration vectors is set as a parameter % agregar a alg
\item RR parameter $\lambda$ was done by cross-validation. %agregar cap timeseries
\end{itemize}

@@

\newpage
The proposal was compared against an optimal online algorithm which is a time varying
VECM (OVECM-OPT). Algorithm ~\ref{alg:OOVECM} shows this modification:

>>>>

Since cointegration vectors represent a long-run relationship between the time
series, they vary little in time. This proposal also includes to obtain new
cointegration vectors only when this long-term relationship changes. This change
is detected by tracking the Mean Absolute Percentage Error (MAPE) defined in the
equation~\eqref{eq:MAPE}, of the last $n$ in-sample forecasts.

@@

\noindent where matrix $\mathbf{B},\mathbf{X},\mathbf{E}$ and $\mathbf{A}$ were
previously defined in equations~\eqref{eq:vecY} ,~\eqref{eq:vecX} ,~\eqref{eq:vecE} and
\eqref{eq:vecA} respectively.

@@

The RR solution $\mathbf{\hat{X}}(t)$ of equation~\eqref{eq:vareqt1} using the
sliding window matrices defined above is:

@@

\begin{itemize}
\item The function \texttt{getJohansen} returns cointegration vectors given by the
Johansen method considering the trace statistic test at 95\% level of
significance. This procedure is called at the first step of the algorithm and
every time in-sample MAPE ($\mathbf{e}$) exceeds threshold error defined (mean\_error).
\item The function \texttt{vecMatrix} returns VECM
matrices $\mathbf{A}(t),\mathbf{B}(t)$ shown in equation~\eqref{eq:notation}. This
method is only required at the first step.
\item The function \texttt{vecMatrixOnline} returns new rows $\mathbf{a}_t^\top$ and
$\mathbf{b}_t^\top$ given new input data $\mathbf{y}_t$.
\item The Solver class (see algorithm \ref{alg:solver}) obtain VECM parameters  
using RR, AAR or OLS methods.
\item The function \texttt{vecMatrixUpdate} updates matrix $\mathbf{A}$ when new
cointegration vectors are required (matrix $\mathbf{B}$ is not affected). Model
parameters matrix $\mathbf{X}$ is also updated later.
\item The number of cointegration vectors is set as a parameter % agregar a alg
\item RR parameter $\lambda$ was done by cross-validation. %agregar cap timeseries
\end{itemize}

@@

\newpage
The proposal was compared against an optimal online algorithm which is a time varying
VECM (OVECM-OPT). Algorithm~\ref{alg:OOVECM} shows this modification:

### 2018-10-22 src/6_Conclusion.tex
correccion despues del examen

>>>>

Financial time series are specially known for presenting high heteroscedasticity, common shocks in volatility could be wrongly interpreted as cointegration. To tackle this problem, a second step of the adaptive VECM algorithm could be implemented to the residuals. A good option could be a GARCH model, for example.
